### 2014-11-13 tex/precise.tex
Finished section 3.1

\section{Generating Summary Heaps}
\label{sec:precise}

In this section we present the rewrite rules for the $\rsym$ relation. 
There are three sets of rewrite rules specific to the summary
heap algorithm: (i) rules to initialize symbolic
references, (ii) field access and write rules, and (iii) rules
to check equality and inequality of references. These rules
form the core of the summary heap algorithm.

@@

During initialization of symbolic references in the $\rgse$ relation,
non-determinitic choices are introduced into the search to
account for initiaization of the reference to $\mathit{null}$, a new instance of the 
object, and to any type compatible object previously
initialized. In contrast, initialization rules for the $\rsym$ relation are 
defined on the bipartite graph consisting of references and locations,
and create a single branch containing a summarization of
all the initializations in a single graph.

@@

The initialization rules are invoked whenever an uninitialized field in a
symbolic reference is accessed. The function $\mathbb{UN}(\cfgnt{L},
\cfgnt{R}, \cfgnt{r}, \cfgnt{f}) = \{\cfgnt{l}\ ...\}$ returns
constraint-location pairs in which field $\cfgnt{f}$ is
uninitialized:
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
where $\mathbb{S}(\phi)$ returns true if $\phi$ is
satisfiable. Intutively, for reference $\cfgnt{r}$, this function constructs
the set, $\theta$ containing all contraint-location pairs that
point to field $\cfgnt{f}$ and $\cfgnt{f}$ points to
$\cfgnt{l}_\mathit{un}$.

@@

The rewrite rule to initialize a heap in the summary heap algorithm
is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $\cfgnt{r}_f$,
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference-location pairs, $(\cfgnt{r}_a\ \cfgnt{l}_a)$, the
field being initialized may alias. A pair $(\cfgnt{r}_a\ \cfgnt{l}_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $\cfgnt{r}_a$ is
part of the input references initialized in the summary heap, (ii)
$\cfgnt{r}_a$ is in the pre-image of $\cfgnt{l}_a$, and (iii) the type 
of $\cfgnt{l}_a$ is the same as the type of the field being initialized.

There are three constraint-location pair sets that correspond to the
non-deterministic choices in GSE: (a) $\theta_\mathit{null}$ contains
a constraint $\cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ paired with the location
$\cfgnt{l}_\mathit{null}$, (b) \sjp{something is not right in this choice}
$\theta_\mathit{new}$ creates a pair with a
constraint representing the fact that $\cfgnt{r}_f$ is not null and it does 
not alias any reference in set $\rho$ under and a newly created location 
$\cfgnt{l}_f$, and (c) 
$\theta_\mathit{alias}$ where for each element in the set $(\cfgnt{r}_a\ \cfgnt{l}_a)
\in \rho$ it creates a constraint stating that $\cfgnt{r}_f$ is not null, points to
$\cfgnt{r}_a$, and does not point to any reference contained in $(\cfgnt{r}_a^\prime\ 
\cfgnt{l}_a^\prime) \in \rho$ such that $\cfgnt{r}_a < \cfgnt{r}_a^\prime$ based on a lexical
ordering of initialized references. Additionally,
$\theta_\mathit{orig}$ allows for the possibility that the field
continues to remain uninitialized since the heap is symbolic.
\sjp{The last statement seems to jump out since we have not been
talking about the heap being symbolic (have we?)} We refer
to initialization in the summary machine as conditional
initialization \sjp{Not sure why we say this - what does it mean
to the reader?}. Finally, the fields of the reference $\cfgnt{r}_f$ are marked
as uninitialized. In the post condition of the rule, the reference
$\cfgnt{r}_f$ points to $\theta$ which contains the union of all the
constraint location pairs sets, and the field $f$ points to the new
reference $\cfgnt{r}_f$ in the summary heap.

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a local
reference. In~\figref{fig:initHeap} $\cfgnt{r}_0^s$ represents a local
reference for the $\cfgt{this}$ reference which has two fields $x$
and $y$ of the same type; reference $\cfgnt{r}_0^s$ points to location
$\cfgnt{l}_0$. Note that when no constraint is specified in the graphs, then
there is an implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The access on $\cfgnt{l}_0.x$ creates a new input reference $\cfgnt{r}_1^i$ in the
summary heap shown in~\figref{fig:initHeap}(b). The reference $\cfgnt{r}_1^1$
points to the $\cfgnt{l}_\mathit{null}$ location under the constraint that
$\cfgnt{r}_1^i$ is null: $\phi_{1a} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null}$. The reference
$\cfgnt{r}_1^i$ points to the location $\cfgnt{l}_1$ under the constraint that $\cfgnt{r}_1^i$
is not $\mathit{null}$ $\phi_{1b} :=\cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null}$. The
location $\cfgnt{l}_1$ represents a fresh location of type $C$ is created on
the heap such that $C$ is type of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$.

>>>>

\section{Generating Summary Heaps}
\label{sec:precise}
There are three sets of rewrite rules specific to the summary heap
algorithm: (i) rules to initialize symbolic references, (ii) rules to complete field
access and field write, and (iii) rules to check equality and
inequality of references. The other rules are common to both $\rsym$
and $\rgse$.

@@

Initialization of uninitialized references in the $\rgse$ relation is
a non-deterministic choice that creates a heap either for
the initialization of the reference to $\mathit{null}$, a new instance of
the object, or to any type compatible object in the input heap (e.g.,
previously initialized by $\rgse$). In contrast, initialization in the
$\rsym$ relation creates a single summary heap containing all of the
possible initialization choices.

@@

The initialization rules, \figref{fig:symInit}, are invoked whenever
an uninitialized field in a symbolic reference is accessed. The
interaction with the solver in the definition of the rules is denoted
by $\mathbb{S}(\phi)$ where it returns true if $\phi$ is
satisfiable. The end rule is active when there is nothing to
initialize. This condition is determined by the function
$\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) =
\{\lp\phi\ \cfgnt{l}\rp\ ...\}$ which returns constraint-location pairs
where the field $\cfgnt{f}$ is uninitialized.
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
For reference $\cfgnt{r}$, this function constructs the set, $\theta$,
of constraint location pairs reachable from $\cfgnt{r}$ on the field
$\cfgnt{f}$ that are valid under the access path from $\cfgnt{r}$.

@@

In~\figref{fig:symInit}, given the unitialized set $\mathbb{UN}$ for
field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexicographical ordering of
uninitialized locations in $\Lambda$ to make initialization
deterministic (the same heap is always initialized in the same
way). 
%The $\mathrm{fresh}_\cfgnt{l}(\cfgnt{C})$ call returns a new location of type $C$. 
The set $\rho$ contains the set of
reference-location pairs from the input heap that are potential
alias. Minimizing over the set of references in the pre-image ensures
that any pair in $\rho$ is the reference $\cfgnt{r}_a$ that birthed
the location $\cfgnt{l}_a$ in some prior initialization (i.e., it was
$\cfgnt{r}_f$ and $\cfgnt{l}_f$ at one time). Aliasing constraint
in the new summary heap reason over these original creators; otherwise aliasing and local value set
computation breaks down.  The reference partitions with their
monotonicity make this determination possible.

There are three sets correspond to the non-deterministic choices in
GSE with a fourth that preserves the pre-initialized heap structure
(the constraints in each preserve the access path to the location):
(i) $\theta_\mathit{null}$ asserts the condition under which
$\cfgnt{l}_\mathit{null}$ is possible (any time the solver assigns
values to reference variables in such a way that $\cfgnt{r}_f =
\cfgnt{r}_\mathit{null}$ holds as does access path $\phi_x$). (ii)
$\theta_\mathit{new}$, meaning $\cfgnt{r}_f$ refers to the fresh
location, occurs when $\cfgnt{r}_f$ is not new or any other alias. (iii) any member of $\theta_\mathit{alias}$ restricts $\cfgnt{r}_f$ to
only alias one member of $\rho$ and not alias any member of $\rho$
that was initialized earlier than the current choice (a later
initialization may indeed alias the current alias choice in a valid
heap but the current choice cannot alias both the choice and a
previous initialization). And (iv) $\theta_\mathit{orig}$  implements conditional initialization to preserve the
original heap structure. This property is needed for the bisumulation. 
These sets are added into the heap on the field after initializing the fields in the new location to $\cfgnt{l}_\mathit{un}$.

\begin{comment} where for each element in the set $(\cfgnt{r}_a\ \cfgnt{l}_a)
\in \rho$ it creates a constraint stating that $\cfgnt{r}_f$ is not null, points to
$\cfgnt{r}_a$, and does not point to any reference contained in $(\cfgnt{r}_a^\prime\ 
\cfgnt{l}_a^\prime) \in \rho$ such that $\cfgnt{r}_a < \cfgnt{r}_a^\prime$ based on a lexical
ordering of initialized references. Additionally,
$\theta_\mathit{orig}$ allows for the possibility that the field
continues to remain uninitialized since the heap is symbolic.
\sjp{The last statement seems to jump out since we have not been
talking about the heap being symbolic (have we?)} We refer
to initialization in the summary machine as conditional
initialization \sjp{Not sure why we say this - what does it mean
to the reader?}. Finally, the fields of the reference $\cfgnt{r}_f$ are marked
as uninitialized. In the post condition of the rule, the reference
$\cfgnt{r}_f$ points to $\theta$ which contains the union of all the
constraint location pairs sets, and the field $f$ points to the new
reference $\cfgnt{r}_f$ in the summary heap.
\end{comment}

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a reference from the stack partition.
 In~\figref{fig:initHeap} $\cfgnt{r}_0^s$ represents a stack
reference for the $\cfgt{this}$ variable which has two fields $x$
and $y$ of the same type; reference $\cfgnt{r}_0^s$ points to location
$\cfgnt{l}_0$. Note that when no constraint is specified in the graphs, then
there is an implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The access on $\cfgnt{l}_0.x$ creates a new input heap reference $\cfgnt{r}_1^i$ in the
summary heap shown in~\figref{fig:initHeap}(b). The reference $\cfgnt{r}_1^1$
points to the $\cfgnt{l}_\mathit{null}$ location under the constraint that
$\cfgnt{r}_1^i$ is null: $\phi_{1a} := \cfgnt{r}_1^i = \cfgnt{r}_\mathit{null}$. The reference
$\cfgnt{r}_1^i$ points to the location $\cfgnt{l}_1$ under the constraint that $\cfgnt{r}_1^i$
is not $\mathit{null}$ $\phi_{1b} :=\cfgnt{r}_1^i \neq \cfgnt{r}_\mathit{null}$. The
location $\cfgnt{l}_1$ represents a fresh location of type $C$ is created on
the heap such that $C$ is type of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$.
